On The Impact of Atoms of Confusion in JavaScript Code

### 2021-10-20 3_Method.tex
minor change
revisao da secao de resultados e discussao

\begin{enumerate}[(RQ1)]
\item \rqa 
\item \rqb
\item \rqc  
\item \rqd
\end{enumerate}

@@

In the first study, we explore the impact of atom candidates on
understanding JavaScript code (goal of research question RQ1).	
To this end, we investigate whether or not code snippets that
contain atom candidates (a) produce a higher
misunderstanding rate when programmers
try to predict their outcome and (b) require
programmers to take longer to predict their output.
Since JavaScript and C have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
Our supplementary material discusses and shows examples of the atoms we consider in our research.

@@

After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The sizes of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

>>>>

\begin{enumerate}[(RQ1)]
\item \rqa 
\item \rqb
% \item \rqc  
\item \rqd
\end{enumerate}

@@

In the first study, we explore the impact of atom candidates on
understanding JavaScript code (goal of research question RQ1).	
To this end, we investigate whether or not code snippets that
contain atom candidates (a) produce a higher
misunderstanding rate when programmers
try to predict their outcome and (b) require
programmers to take longer to predict their output.
Since JavaScript and \clang have some constructs in common (Section~\ref{back}), we first selected a set of 9+1 atom candidates: 9 atoms of confusion already validated~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} for the \clang language that also exist in JavaScript programs + 1 atom candidate that is specific of the JavaScript language (Automatic Semicolon Insertion).
Our supplementary material discusses and shows examples of the atoms we consider in our research.

@@

After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

### 2021-10-20 4_Result.tex
revisao da secao de resultados e discussao

\subsubsection*{Exploratory Data Analysis}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions. We do not consider cases where the participants made mistakes. Seven atom candidates required more time from participants to predict a correct response. For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased  
29.06\% and 38.19\% for the clean versions.

@@

\
% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}

\vspace*{-0.5cm}
\section{Results of the Interview Study}
\label{sec:interview-results}

@@

In this section, we present the results of the 
interviews with the 15 practitioners who agreed to participate in the study. 
% A total of 15 practitioners took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in our research. We first presented them pairs of clean and confusing code snippets and then asked them to discuss their preferences towards one version. We did not indicate in any way whether any version was assumed to be confusing or not. 
% ask the participants to discuss their preferences (i.e., code with or without atom).

@@

\begin{mq}
\emph{``When I am programming, I write code with the ternary operator, [...], but, to be honest, I still think that the 
[the code using the \rhs version] 
is easier to understand''}.
\end{mq}

@@

When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example personal experience that might motivate one to avoid writing code using this construct:

\begin{mq}
  \emph{``This one is interesting, because I have written code that looks like the left-hand side [Confusing version], and my colleagues complained that it was difficult to understand. Nowadays I prefer to write code using the version on the right-hand side [Clean]''}.
\end{mq}

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript. Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern. We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).

@@

\todo[inline]{Seria legal contrastar esses achados com o que foi reportado anteriormente}

>>>>

\subsubsection*{Exploratory Data Analysis}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both confusing and clean versions. We do not consider cases where the participants made mistakes. Seven atom candidates required more time from participants to predict a correct response. For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for the Arithmetic as Logic and the Pre-Increment atoms, the time to give a correct answer increased  
29.06\% and 38.19\% for the clean versions---we did not confirm these
atoms as atoms of confusion in the previous section.

@@

% \rb{acho que podemos melhorar a apresenta\c c\~{a}o dessas tabelas, talvez usando booktabs.}

Next, we summarize the results of our second study (interview with
practitioners), which
we use to contrast the findings discussed here with the opinion
of software developers about their preferences regarding the confusing
or clean versions of the code snippets we used in our experiment.

\section{Results of the Interview Study}
\label{sec:interview-results}

@@

%In this section, we present the results of the 
% interviews with the 15 practitioners who agreed to participate in the study. 
A total of 15 practitioners took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion about the \na atom candidates we explored in our research. We first presented them pairs of clean and confusing code snippets and then asked them to discuss their preferences towards one version. We did not indicate in any way whether any version was assumed to be confusing or not. 
% ask the participants to discuss their preferences (i.e., code with or without atom).

@@

\begin{mq}
\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that the 
[the code using the \rhs version] 
is easier to understand''}.
\end{mq}

@@

When analyzing the Logic as Control Flow atom candidate, one of the interviewees gave an example of personal experience that might motivate one to avoid writing code using this construct:

\begin{mq}
  \emph{``This one is interesting, because I have written code that looks like the left-hand side [confusing version], and my colleagues complained that it was difficult to understand. Nowadays I prefer to write code using the version on the right-hand side [clean version]''}.
\end{mq}

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript. Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern. We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).  As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code. The following section presents the results of our mining software repository effort, whose goal is to reveal how frequent atom candidates are in open source code.

@@

### 2021-10-20 discussion.tex
revisao da secao de resultados e discussao

Our work has several implications.
First, it adds external validity to the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang and \cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, Logic as
Control Flow, and Ternary Operator seem to make 
JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
In comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
our study
led to some differences in the effect size
of the atom candidates.
Altogether, we answer our first research question
\emph{\rqa}:

@@

The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes. For instance, most of the participants prefer the version of the code with the Ternary Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase. The participants of the interview study also
mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-based inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion.
In summary, the results of the second study (interviews) allow
us to answer the second (\rqb) and third research questions
(\emph{\rqc}).

\begin{mh}
  {\bf Answer to RQ2:} The qualitative analysis of the
  interviews supports the results of the first study,
  indicating that atoms of confusion compromise
  source code understanding. 
\end{mh}

\begin{mh}
  {\bf Answer to RQ3:} The qualitative analysis of the
  interviews suggests that specific JavaScript constructs might also correspond to atoms of
  confusion, including prototype inheritance and
  nested callbacks. 
\end{mh}

The results of the third study (mining open source
Java script repositories) provides evidence that,
although atoms of confusion compromise program
comprehension, they frequently appear in open
source JavaScript projects. In particular,
seven, out of 10 atoms considered
  in our study, appear in more than 50\% of
the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
allows us to answer the fourth research
question (\emph{\rqd}).

\begin{mh}
  {\bf Answer to RQ4:} The MSR study reveals that
  the atoms of confusion explored in our research
  appear frequently in practice. 
\end{mh}

>>>>

Our work has several implications.
First, it adds external validity to the
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17},
which investigates
the impact of atom candidates on
understanding \clang,\cpplang code. That is,
similarly to their work, the atom candidates
Comma Operator, Post/Pre Increment, Omitted Curly Braces,
Assignment as Value, Implicit Predicate, and Logic as
Control Flow seem to make 
JavaScript code harder to understand. For five of them, the difference is statistically significant, with a large effect size for four atoms. Our results also 
refute the hypothesis that Arithmetic as Logic is an atom of confusion (i.e., a source of misunderstanding).
In comparison to the original
work of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, 
our study
led to some differences in the effect size
of the atom candidates.
Altogether, we answer our first research question
\emph{\rqa}:

@@

The results of the interview study complement the understanding of atoms of confusion because the participants make clear the existence of a trade-off between code comprehension and other quality attributes. For instance, most of the participants prefer the version of the code with the Conditional Operator, even though they agree that its use might contribute to the misunderstanding of JavaScript code, particularly when novices are maintaining the codebase. The participants of the interview study also
mentioned other possible sources of misunderstanding in JavaScript,
including the use of prototype-based inheritance and nested call-backs (as discussed in Section~\ref{sec:interview-results}). Other JavaScript atom candidates include
Object Destructuring, Array Spread, Object Spread, and Type Conversion.
In summary, the results of the second study (interviews) allow
us to answer the second (\rqb) research question.

\begin{mh}
  {\bf Answer to RQ2:} The qualitative analysis of the
  interviews supports the results of the first study,
  since JavaScript developers most often agree that atoms of confusion compromise
  source code understanding. 
\end{mh}

%% \begin{mh}
%%   {\bf Answer to RQ3:} The qualitative analysis of the
%%   interviews suggests that specific JavaScript constructs might also correspond to atoms of
%%   confusion, including prototype inheritance and
%%   nested callbacks. 
%% \end{mh}

The results of the third study (mining open source
JavaScript repositories) provides evidence that,
although atoms of confusion compromise program
comprehension, they frequently appear in open
source JavaScript projects. In particular,
seven, out of 10 atoms considered
  in our study, appear in more than 50\% of
the 72 projects we analyzed. Furthermore, at least two of them are used intensively, more than once for every 200 lines of code. In summary, the third study
allows us to answer the fourth research
question (\emph{\rqd}).

\begin{mh}
  {\bf Answer to RQ3:} The MSR study reveals that
  the several atom candidates explored in our research
  appear frequently in practice, and cleaning up the use of 
  Post-Increment/Decrement and the Automatic Semicolon Insertion
  might improve the readabiliy of JavaScript code substantially. 
(\emph{\rqc}) 
\end{mh}
